---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Riccardo Fusaroli"
date: "October 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia, focusing on pitch.
In the course of this assignment we will use them to try to automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.

### Question 1
Build a logistic regression to see whether you can diagnose schizophrenia from pitch range only.

```{r}
# Some packgages
pacman::p_load(knitr,lme4,caret,e1071,pROC,dplyr)

# Read in data
data = read.csv("final_rqa.csv")

data$participant = as.factor(data$participant)
data$study = as.factor(data$study)

# Scale everything
data <- data %>% mutate_each_(funs(scale(.) %>% as.vector), 
                             vars=c("mean","stdDev","range","median","InterquartileRange","MeanAbsoluteDeviation","coefficientOfVariation","delay","radius","embed","rqa_REC","rqa_DET","rqa_maxL","rqa_L","rqa_ENTR","rqa_TT","rqa_LAM"))

# Make an identical original dataframe
d = data

# Visualize
ggplot(data, aes(diagnosis, range)) +
  geom_boxplot() +
  ggtitle("Pitch range in controls and Schizophrenia patients")

# Predict schizophrenia from pitch range only
model1 = glm(diagnosis ~ range, family = binomial, data)
inv(summary(model1))

## not the best results
# log odds tho
exp(0.0481705)

#How many skisos and controls do we have
table(data$diagnosis)
m = data %>%
  group_by(participant) %>%
  count(diagnosis == "control")

colnames(m) = c("participant", "diagnosis", "n")
table(m$diagnosis)
#57 controls and 59 skiso

## 1.04935 control:skiso (but check pls)
``` 

Calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve) on a logistic regression using the full dataset. Don't forget the random effects!

```{r}
# Start off with a glmer model (because you introduce random effects you can't really use glm anymore
model2 = glmer(diagnosis ~ range + (1|study), family = binomial, data)
summary(model2)
library(boot);library(lme4)
#baseline
inv.logit(fixef(model2)[[1]])
#range
inv.logit(fixef(model2)[[2]])
##rescale variables - not really an issue

# Predictions in a new column, made based on model2
data$PredictionsPerc = predict(model2)

# Get the mean and range to start making guesses on the threshold
mean(data$PredictionsPerc)
range(data$PredictionsPerc)

# Set the treshold and run the confusion matrix, adjust threshold based on matrix (want to be high in sensitivity and low in specificity)
data$Predictions[data$PredictionsPerc > 0] = "control"
data$Predictions[data$PredictionsPerc <= 0] = "schizophrenia"
theMatrix = confusionMatrix(data = data$Predictions, reference = data$diagnosis, positive = "schizophrenia") 
theMatrix

# Get the accuracy, sensitivity and specificity, PPV and NPV
(accuracy = theMatrix$overall["Accuracy"])
(sensitivity = theMatrix$byClass["Sensitivity"])
(specificity = theMatrix$byClass["Specificity"])
(PPV = theMatrix$byClass["Pos Pred Value"])
(NPV = theMatrix$byClass["Neg Pred Value"])

str(theMatrix)

# ROC curve
rocCurve = roc(response = data$diagnosis, predictor = data$PredictionsPerc)
# Area under curve
auc(rocCurve) ## 0.5752 under curve
auc(roc(response = data$diagnosis, predictor = data$PredictionsPerc))
# Confidence intervals
ci(rocCurve)
# Plot the curve, see where the specificity-sensitivity threshold should be, incorporate that in the above code (aim for 60% sensitivity)
plot(rocCurve, legacy.axes = T)
```
accuracy = 0.44, 
sensitivity = 0.31, 
# samples with the event and predicted to have the event/samples having the event
specificity = 0.57, 
# samples without event and predicted as nonevents/samples without the event
PPV = 0.41, 
NPV = 0.46, 
ROC curve

Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures.

smart solution: cvms package, write list of all models u want to test, tests all the models

```{r}
### CROSS-VALIDATED VERSION OF THE MODEL
#Create ID as numbers for use in folding
d$fold_id = as.numeric(d$participant)
# Create folds
folds = createFolds(unique(d$fold_id), 5)

# To save results in a dataframe
Results_range = as.data.frame(NULL)

# Create empty lists
acc_test = NULL
sensi_test = NULL
speci_test = NULL
PPV_test = NULL
NPV_test = NULL

rocC = NULL
AUC_test = NULL
CI_test = NULL

n = 1

for(i in folds) {
  #select data
  dtraint = subset(d, !fold_id %in% i)
  dtestt = subset(d, fold_id %in% i)
  #train model
  model2 = glmer(diagnosis ~ range + (1|study), family = "binomial", dtraint)

  # Predictions in a new column, made based on model2
  dtestt$PredictionsPerc = predict(model2, dtestt, allow.new.levels = T)
  
  # Set the threshold and run the confusion matrix, adjust threshold based on matrix (want to be high in sensitivity and low in specificity)
  dtestt$Predictions[dtestt$PredictionsPerc > 0] = "control"
  dtestt$Predictions[dtestt$PredictionsPerc <= 0] = "schizophrenia"
  
  CM = confusionMatrix(data = dtestt$Predictions, reference = dtestt$diagnosis, positive = "schizophrenia")

  acc_test[n] = CM$overall["Accuracy"]
  sensi_test[n] = CM$byClass["Sensitivity"]
  speci_test[n] = CM$byClass["Specificity"]
  PPV_test[n] = CM$byClass["Pos Pred Value"]
  NPV_test[n] = CM$byClass["Neg Pred Value"]
  
  # ROC curve, area under curve and confidence intervals
  rocC = roc(response = dtestt$diagnosis, predictor = dtestt$PredictionsPerc)
  AUC_test[n] = auc(rocC)
  CI_test[n] = ci(rocC)
  
  #save the performance -- add to list c()
  n = n+1
  
}

results_range = data.frame(Model = "diagnosis ~ range + (1|study)",
                     Accuracy = mean(acc_test, na.rm= T),
                     Sensitivity = mean(sensi_test, na.rm= T),
                     Specificity = mean(speci_test, na.rm= T),
                     PPV = mean(PPV_test, na.rm= T),
                     NPV = mean(NPV_test, na.rm= T),
                     AUC = mean(AUC_test, na.rm= T),
                     CI = mean(CI_test, na.rm= T))

```


N.B. the predict() function generates probabilities (the full scale between 0 and 1). A probability > .5 indicates a choice of 1, below a choice of 0.
N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one datase, so to calculate overall performance.
N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?

### Question 2

Which single predictor is the best predictor of diagnosis?

start cross-validating to not overfit with just a regression, area under curve %

```{r}
library(caret)
data$participant = as.numeric(data$participant)

# List of all the models that will be tested
models = c("diagnosis ~ mean + (1|study)", 
       "diagnosis ~ stdDev + (1|study)",
       "diagnosis ~ range + (1|study)",
       "diagnosis ~ median + (1|study)",
       "diagnosis ~ InterquartileRange + (1|study)",
       "diagnosis ~ MeanAbsoluteDeviation + (1|study)",
       "diagnosis ~ coefficientOfVariation + (1|study)",
       "diagnosis ~ delay + (1|study)",
       "diagnosis ~ radius + (1|study)",
       "diagnosis ~ embed + (1|study)",
       "diagnosis ~ rqa_REC + (1|study)",
       "diagnosis ~ rqa_DET + (1|study)",
       "diagnosis ~ rqa_maxL + (1|study)",
       "diagnosis ~ rqa_L + (1|study)",
       "diagnosis ~ rqa_ENTR + (1|study)",
       "diagnosis ~ rqa_LAM + (1|study)",
       "diagnosis ~ rqa_TT + (1|study)"
       )

# To save results in a dataframe
Results = as.data.frame(NULL)

# Loop for ALL MODELS
for (model in models){

# Create folds
folds = createFolds(unique(data$participant), 5)
#d = data[data$participant %in% folds]
## why do i have this?

acc_test = NULL
sensi_test = NULL
speci_test = NULL
PPV_test = NULL
NPV_test = NULL

roc_test = NULL
AUC_test = NULL
CI_test = NULL

n = 1

for(i in folds) {
  #select data
  dtraint = subset(data, !(participant %in% i))
  dtestt = subset(data, participant %in% i)
  #train model
  model2 = glmer(model, dtraint, family = "binomial")
  
  # Predictions in a new column, made based on model2
  dtestt$PredictionsPerc[n] = predict(model2)
  
  # Set the treshold and run the confusion matrix, adjust threshold based on matrix (want to be high in sensitivity and low in specificity)
  dtestt$Predictions[dtestt$PredictionsPerc > 0][n] = "control"
  dtestt$Predictions[dtestt$PredictionsPerc <= 0][n] = "schizophrenia"
  
  CM = confusionMatrix(data = dtestt$Predictions, reference = dtestt$diagnosis, positive = "schizophrenia")

  acc_test[n] = CM$overall["Accuracy"]
  sensi_test[n] = CM$byClass["Sensitivity"]
  speci_test[n] = CM$byClass["Specificity"]
  PPV_test[n] = CM$byClass["Pos Pred Value"]
  NPV_test[n] = CM$byClass["Neg Pred Value"]
  
  # ROC curve, area under curve and confidence intervals
  rocC = roc(response = dtestt$diagnosis, predictor = dtestt$PredictionsPerc)
  AUC_test[n] = auc(rocC)
  CI_test[n] = ci(rocC)
  
  #save the performance -- add to list c()
  n = n+1
}

# Add results to a row to add to dataframe
add_row = data.frame(Model = model,
                     Accuracy = mean(acc_test, nar.rm = T),
                     Sensitivity = mean(sensi_test, nar.rm = T),
                     Specificity = mean(speci_test, nar.rm = T),
                     PPV = mean(PPV_test, nar.rm = T),
                     NPV = mean(NPV_test, nar.rm = T),
                     AUC = mean(AUC_test, nar.rm = T),
                     CI = mean(CI_test, nar.rm = T))

# Bind final dataframe with all results
Results = rbind(Results, add_row)

}
```


### Question 3

Now it's time to go wild! Use all (voice-related) variables and interactions you can think of. Compare models and select the best performing model you can find.

Remember:
- Cross-validation or AIC are crucial to build the best model!
- After choosing the model, train it on all the data you have
- Save the model: save(modelName, file = "BestModelForever.rda")
- Create a Markdown that can: a) extract the features from new pitch files (basically your previous markdown), b) load your model (e.g. load("BestModelForever.rda")), and c) predict the diagnosis in the new dataframe.
Send it to Celine and Riccardo by Monday (so they'll have time to run it before class)-

```{r}
library(dplyr)

# List of all the models that will be tested
goodmodels = c("diagnosis ~ mean + stdDev + range + median + InterquartileRange + MeanAbsoluteDeviation + coefficientOfVariation + delay + radius + embed + rqa_REC + rqa_DET + rqa_L + rqa_ENTR + rqa_LAM + rqa_TT + (1|study)", ## All
       "diagnosis ~ mean + stdDev + range + median + embed + rqa_DET + rqa_maxL + rqa_ENTR + (1|study)", # 50% best
       "diagnosis ~ range + median + embed + rqa_DET + rqa_maxL + rqa_ENTR + (1|study)", #40% best
       "diagnosis ~ median*range*mean + rqa_DET + (1|study)", #assume interactions
       "diagnosis ~ mean + rqa_DET*rqa_REC*rqa_maxL*rqa_ENTR + (1|study)", #
       "diagnosis ~ rqa_DET*rqa_L + (1|study)",
       "diagnosis ~ median*range*mean + rqa_DET*rqa_REC*rqa_maxL*rqa_ENTR + (1|study)"
       )
# To save results in a dataframe
BestModel = as.data.frame(NULL)

# Loop for ALL MODELS
for (model in goodmodels){

# Create folds
folds = createFolds(unique(data$participant), 5)
#d = data[data$participant %in% folds]
## why do i have this?

acc_test = NULL
sensi_test = NULL
speci_test = NULL
PPV_test = NULL
NPV_test = NULL

roc_test = NULL
AUC_test = NULL
CI_test = NULL

n = 1

for(i in folds) {
  #select data
  dtraint = subset(data, !(participant %in% i))
  dtestt = subset(data, participant %in% i)
  #train model
  model2 = glmer(model, dtraint, family = "binomial")
  
  # Predictions in a new column, made based on model2
  dtestt$PredictionsPerc[n] = predict(model2)
  
  # Set the treshold and run the confusion matrix, adjust threshold based on matrix (want to be high in sensitivity and low in specificity)
  dtestt$Predictions[dtestt$PredictionsPerc > 0][n] = "control"
  dtestt$Predictions[dtestt$PredictionsPerc <= 0][n] = "schizophrenia"
  
  CM = confusionMatrix(data = dtestt$Predictions, reference = dtestt$diagnosis, positive = "schizophrenia")

  acc_test[n] = CM$overall["Accuracy"]
  sensi_test[n] = CM$byClass["Sensitivity"]
  speci_test[n] = CM$byClass["Specificity"]
  PPV_test[n] = CM$byClass["Pos Pred Value"]
  NPV_test[n] = CM$byClass["Neg Pred Value"]
  
  # ROC curve, area under curve and confidence intervals
  rocC = roc(response = dtestt$diagnosis, predictor = dtestt$PredictionsPerc)
  AUC_test[n] = auc(rocC)
  CI_test[n] = ci(rocC)
  
  #save the performance -- add to list c()
  n = n+1
}

# Add results to a row to add to dataframe
add_row = data.frame(Model = model,
                     Accuracy = mean(acc_test, na.rm = T),
                     Sensitivity = mean(sensi_test, na.rm = T),
                     Specificity = mean(speci_test, na.rm = T),
                     PPV = mean(PPV_test, na.rm = T),
                     NPV = mean(NPV_test, na.rm = T),
                     AUC = mean(AUC_test, na.rm = T),
                     CI = mean(CI_test, na.rm = T))

# Bind final dataframe with all results
BestModel = rbind(BestModel, add_row)

}
```


### Question 4: Report the results

METHODS SECTION: how did you analyse the data?

RESULTS SECTION: can you diagnose schizophrenia based on voice? which features are used? Comment on the difference between the different performance measures.

### Bonus question 5

You have some additional bonus data involving speech rate, pauses, etc. Include them in your analysis. Do they improve classification?

### Bonus question 6

Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them.

(same code, just replace the log reg part)
